Polymers for Near-field Electrospinning with Spatial Control

### 2019-09-11 NFES_Review/article.tex
article progress

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Feature enabled:
%toc: yes
%pagenum: yes
%text-layout: twocolumn
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\def\ps@pprintTitle{\save@ps@pprintTitle\gdef\@oddfoot{\footnotesize\hspace*{.5\textwidth}\thepage\itshape \null\hfill\today}}
\makeatother
          \makeatletter\@ifundefined{tableofcontents}{\usepackage{typeset-custom-toc}}{}\makeatother
\usepackage{etoolbox}

@@

\begin{document}

@@

\section{Introduction}
Even though electrospinning is an old invention \unskip~\cite{527120:12073288}, it is currently a trending topic among researchers \unskip~\cite{527120:12073453,527120:12073495,527120:12073496}. One of the reasons electrospinning is to be studied is its potential to fabricate polymer nano-fibers from a variety of polymers. The technique allows the production of thin continuous fibers with ease, with diameters down to 3 $nm $ in some cases, which is something difficult to achieve by other techniques. Furthermore, the basic setup can be modified with ease to fabricate different fibers with diversified functionalities with different materials. The produced fibers can be aligned or unaligned. Besides, the electrospinning equipment is inexpensive and of small size, compared to the equipment of standard spinning techniques. On the other hand, the understanding of the electrospinning process has improved in the last years \unskip~\cite{527120:12073538}.

\subsection{Types of electrospinning }The main components of the electrospinning technique are the fluid control unit (e.g. syringe pump) and a voltage power supply. The process also requires a target electrode or combination of electrodes on which the fibers can be collected.  Figure~\ref{f-fe28447572e9} describes a typical near-field electrospinning set-up \unskip~\cite{527120:12073538}. Two sub-techniques can be derived from electrospinning depending on the distance between the dispensing electrode and the collector. The process in which the electrospun jet can be controlled near the tip is called NFES or near-field electrospinning \unskip~\cite{527120:12033655}. Moreover, if the distance between the collector and the dispensing needle is greater, the configuration is known as FFES or far-field electrospinning \unskip~\cite{527120:12073581}.

\subsubsection{High voltage power supply}Direct Current \mbox{}\protect\newline Alternating Current

\subsubsection{Polymer reservoir}Polymer melt \mbox{}\protect\newline Polymer solution

\subsubsection{Stretching forces}Centrifugal force \mbox{}\protect\newline Blowing forces \mbox{}\protect\newline Electric force \mbox{}\protect\newline Microfluidic forces \mbox{}\protect\newline Mechanical force

\subsubsection{Polymer yielding}Atomization (particles) \mbox{}\protect\newline Spraying (droplets) \mbox{}\protect\newline Spinning (fibers)

\subsubsection{Dispensing nozzle}Coaxial / Monoaxial \mbox{}\protect\newline Multinozzle

\subsubsection{Nozzle-to-substrate distance}Far Field Spinning \mbox{}\protect\newline Near Field Spinning

@@

\section{NFES Variants}
[SECTION UNDERWORK]

\bgroup
\fixFloatSize{images/69d16de4-c7d7-4000-b4c8-93a40c79692b-uelectrospinningvariants.jpg}
\begin{figure*}[!htbp]
\centering \makeatletter\IfFileExists{images/69d16de4-c7d7-4000-b4c8-93a40c79692b-uelectrospinningvariants.jpg}{\includegraphics{images/69d16de4-c7d7-4000-b4c8-93a40c79692b-uelectrospinningvariants.jpg}}{}
\makeatother 
\caption{{}}
\label{f-f076e2b74317}
\end{figure*}
\egroup

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Feature enabled:
%toc: yes
%bookmark: yes
%pagenum: yes
%text-layout: twocolumn
%endfloat: placeins
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\makeatletter
\@ifpackageloaded{hyperref}{\def\alreadybookmark{}}{\usepackage[hidelinks]{hyperref}}
\@ifpackageloaded{bookmark}{}{\usepackage{bookmark}}
\makeatother
\ifdefined\alreadybookmark\def\pdfbookmark[#1]#2#3{}\fi
\makeatletter
\def\ps@pprintTitle{\save@ps@pprintTitle\gdef\@oddfoot{\footnotesize\hspace*{.5\textwidth}\thepage\itshape \null\hfill\today}}
\makeatother
          \makeatletter\@ifundefined{tableofcontents}{\usepackage{typeset-custom-toc}}{}\makeatother
\usepackage[section]{placeins}

\usepackage{etoolbox}

@@

\begin{document}
\pdfbookmark[title]{Polymers for Near-field Electrospinning with Spatial Control}{title}

@@

\section{Types of electrospinning (Classified by process properties)}
~Even though electrospinning is an old invention \unskip~\cite{527120:12073288}, it is currently a trending topic among researchers \unskip~\cite{527120:12073453,527120:12073495,527120:12073496}. One of the reasons electrospinning is to be studied is its potential to fabricate polymer nano fibers from a variety of polymers. The technique allows the production of thin continuous fibers with ease, with diameters down to 3 $nm $ in some cases, which is something difficult to achieve by other techniques. Furthermore, the basic setup can be modified with ease to fabricate different fibers with diversified functionalities with different materials. The produced fibers can be aligned or unaligned. Besides, the electrospinning equipment is inexpensive and of small size, compared to the equipment of standard spinning techniques\unskip~\cite{527120:12073538}. On the other hand, the understanding of the electrospinning process has improved in the last years.

Current literature dictates the typical spinning setup is comprised by three main components: a syringe needle, a fiber collector, and some way to dispense the fibers from the needle to the collector. The spinning process is an electrohydro-dynamic (EHD) technique that yields continuous polymer fibers. Other EHD techniques are spraying and atomization which produce polymer droplets and polymer particles respectively.

\bgroup
\fixFloatSize{images/f10bcbc9-719b-4a8d-8a38-930cc6ea2d0f-uimg_ehdprocesses.jpg}
\begin{figure*}[!htbp]
\centering \makeatletter\IfFileExists{images/f10bcbc9-719b-4a8d-8a38-930cc6ea2d0f-uimg_ehdprocesses.jpg}{\includegraphics{images/f10bcbc9-719b-4a8d-8a38-930cc6ea2d0f-uimg_ehdprocesses.jpg}}{}
\makeatother 
\caption{{Electrohydro-dynamic techniques}}
\label{f-02e0e3cf88d6}
\end{figure*}
\egroup
In electrospinning the fibers are deposited by an electrical potential difference between the syringe needle and the collector. The supplied polymer (typically a polymer solution) is administrated at a constant rate to create and maintain a polymer drop at the dispensing nozzle. A high voltage (usually DC) is applied between the polymer feed and the collector. As the electric field increases, the polymer drop (held by its surface tension) is then deformed at the tip of the syringe needle to form a conical shape known as Taylor cone. When the electric force overcomes the surface tension force a polymer jet is ejected from the tip of the Taylor cone. As the polymer jet leaves the nozzle, it accelerates and stretches while traveling to the fiber collector. The fiber finally develops with the complete solvent evaporation.

\subsection{High voltage power supply: Direct Current \& Alternating Current}Direct current (DC) is typically used in electrospinning with the electrons flowing in one direction. Alternate current (AC) implementations are also studied as the AC creates a change in the direction of the current flow. Kessick et al.\unskip~\cite{527120:13444381} demonstrated the implementation of AC power supplies in the production of polymer fibers.

The AC electrospinning setup is similar to that for the DC variant. AC electrospinning apparatus do not require a grounded collector as the current alternates. In AC, the produced fibers are prone to carry an electric charge, while those generated shortly after have an opposite charge. The difference in charges lead the fibers to discharge on each other, creating an aerogel plume of fibers. The optimal AC frequency depends on the materials used and is typically within of 50 $Hz $ {\textendash} 1 $kHz $. \unskip~\cite{527120:13443405}

The AC technique has been studied for drug loaded related applications. Balogh et al. \unskip~\cite{527120:13445177} compared fibers fabricated by DC and AC spinning techniques. They produced fibers by using three polymers: Eudragit EPO (cationic copolymer), Eudragit L100-55 (anionic polymer) and poly(vinyl pyrrolidone) (PVP) (neutral polymer). Balogh et al. discovered AC and DC electrospinning can produce fibers with all three polymers. However, The AC process allowed the implementation of faster flow rates than in the DC setup. The DC electrospinning technique generated fibers with a maximum flow rate of 5 $ml/h $; on the other hand, the AC setup allowed an increase in flow rate up to 40 $ml/h $. AC electrospinning approach is as effective as the DC technique in producing drug delivery systems.

\subsection{Polymer reservoir: Polymer melt \& Polymer solution}Another electrospinning classification is the melt process. As Brown et al. \unskip~\cite{527120:13445499} discussed, the polymer melt is equivalent to the polymer solution electrospinning (in place of a polymer solution a melt is used). The use of a polymer melt increases the complexity of the process, because the nozzle syringe and spinneret required to be heated to maintain the polymer in a liquid state. The fibers produced in melt spinning are typically found to have larger diameters than those from the polymer solutions due to the higher viscosity of a polymer melt than its solution. The apparatus used by Brown et al. \unskip~\cite{527120:13445499} is depicted in .

[TODO: draw an apparatus used for melt electrospinning].

Despite the added complexity and thicker diameters, melt electrospinning  gets around the need to handle volatile solvents, making the process safer to be performed on larger scales, On the other hand, this technique gets rid of any solvent contamination.

The first report of a melt electrospun drug delivery system came from Nagy et al. \unskip~\cite{527120:13445555}, who prepared fibers by melt electrospinning of Eudragit EPO with carvedilol. The drug and polymer were melted and mixed to form a homogeneous solid mixture prior to spinning. The melt-spun fibers reached diameters of 5{\textendash}30 $\mu m $, compared to 300{\textendash}1000 $nm $ diameters produced from solution-spun fibers.

This work has been built on to blend plasticisers with the polymer Eudragit EPO and carvedilol active ingredient.\unskip~\cite{527120:13445752} The plasticisers Triacetin, Tween 80 and Polyethylene Glycol were investigated by Attila et al. in order to reduce the melting point of the polymer-drug mixture. The temperature drop is desirable to minimize the occurring drug degradation.

Lian and Meng\unskip~\cite{527120:13445754} performed a comparison of poly(\ensuremath{\varepsilon }-caprolactone) (PCL) fibers fabricated by the melt and solution electrospinning techniques. Lian and Meng\unskip~\cite{527120:13445754} reported that melt spinning is preferable when the polymer presents a low solubility, on the other hand the melt fibers were produced in a slower release rate.  Lian and Men findings feature the solution-spun fibers to have a porous structure. Gernot et al.\unskip~\cite{527120:13534159} demonstrated that submicron-size fibers are possible through melt electrospinning. In their effort, they achieved a precise deposition of PCL fibers with diameters of $817 \pm 165 nm $.

In literature, melt electrospinning has less evidence than the solution approach. However, melt electrospinning arises to be as flexible as its solution counterpart in handling multiple polymers, as reported in McCann's work\unskip~\cite{527120:13534572}. Currently, the melt electrospinning setup is harder to establish but the lack of research on this technique explains its unexplored potential.

\subsection{Stretching forces}

\subsubsection{Electric Field}

\subsubsection{Centrifugal force}The spinning processes require the implementation of a force to break the polymer source into a polymer jet. Centrifugal spinning intends to produce fibers by the use of a rotating polymer source. The centrifugal force generated from typical rotatory speeds above 2000 rpm, results in fiber formation. \unskip~\cite{527120:13535559,527120:13535561}.

The centrifugal force technique is applied to polymer solutions and melts. This approach is used in applications were the precise deposition of the fibers is not relevant and production rate is to be maximized \unskip~\cite{527120:13535894}.  Efforts in centrifugal spinning are focused on drug delivery applications. Zander \unskip~\cite{527120:13535977} fabricated polycaprolactone (PCL) fibers using the solution and melt variants of the centrifugal approach. Zander's fibers were produced with rotatory speed between three and 18 thousand revolutions per minute with $10 \mu m $ in diameter.

On the other hand, PCL and PVP fibers were generated by Amalorpava et al. \unskip~\cite{527120:13536089}.  Amalorpava achieved sub micron/size fiber diameters for drug release purposes and bacteria growth inhibition properties. Literature\unskip~\cite{527120:13536446} has shown that centrifugal approach has a simple setup that promises a large scale fabrication of fibers.

In some cases the centrifugal force implementations and pressurized gyration can be combined with an electric field. The implementation of two stretching forces (centrifugal and electrical forces), can help solvent evaporation\unskip~\cite{527120:13536560}. Centrifugal electrospinning implements the same setup as the standard centrifugal spinning with the addition of a high voltage power supply between the rotating dispensing nozzle and the collector. The combined method has evidence to yield parallel fibers\unskip~\cite{527120:13536841,527120:13536900,527120:13537392,527120:13537393} at a higher rate \unskip~\cite{527120:13536841,527120:13536900} than the standard electrospinning approach.

\subsubsection{Blowing forces}Nano fibers can be produced with the implementation of pressurized gas with a polymer solution. The setup used for blow spinning is similar to the one used in standard electrospinning, where the polymer precursor is dispensed at a controlled rate. Unlike traditional electrospinning, in the solution blow spinning setup the needle nozzle applies pressurized gas to the polymer solution through an outer spinneret\unskip~\cite{527120:13538056}.

[TODO: draw an apparatus used for solution blow spinning or melt blowing] \unskip~\cite{527120:13538056}.

A few studies have explored solution blowing in drug delivery, with the first such work being from Oliveira et al. in 2013. [27] These authors prepared poly(lactic acid) (PLA) fibres loaded with the hormone progesterone, which can be used to regulate the reproductivecycle in livestock. Fibres were produced from solutions with 6\% w/vPLA and between 0 and 8\% w/v progesterone. The PLAis semi-crystalline both before and after processing, while the drug isamorphous post-spinning. The fibres behave very similarly in termsof their release behaviour, regardless of the amount of drug loaded.

A study comparing electrospun and solution-blown fibres of poly(3-hydroxybutyrate-co-3-hydroxyvalerate) loaded with sodium diclofenachas also been reported. [26] The drug-loaded fibres were slightly larger in diameter when generated by electrospinning, and the size uniformity was higher through solution blowing. In general there was a greateramount of burst release seen with the electrospun fibres, but otherwisethere were no clear trends in the drug release data.

Solution-blown fibres have additionally been created loaded with oil extracted from the medicinal plant Copaifera sp., which is often explored for antimicrobial purposes. [28] These materials were constructed from a blend of the polymers PLA and PVP, and were around 1 \ensuremath{\mu}m in diameter. An increased PVP content was found to result in increased antibacterial activity after 24 h.

The melt blowing process has also received some attention in thepharmaceutical setting, and the fibres produced compared with thosefrom both solution and melt electrospinning. [29] Marosi's team generatedformulations from a vinylpyrrolidone{\textendash}vinyl acetate copolymer, employingpoly(ethylene glycol) (PEG) as a plasticiser and carvedilol as a modeldrug. All three methods led to fibres, with the solution electrospun fibres narrowest (at 2 \ensuremath{\mu}m in diameter), followed by the melt-blown (10\ensuremath{\mu}m) and melt-spun (50 \ensuremath{\mu}m) products. Carvedilol was rendered into the amorphous physical form by all three processing techniques, and all theformulations were able to accelerate the drug dissolution process. Themelt-blown and melt electrospun systems led to the fastest release, withalmost identical release profiles, while the solution-electrospun fibresfreed their drug cargo somewhat more slowly.

A variant of the solution-blowing technique has been applied tothe processing of living cells (in this setting it has been referred to asbiothreading). [30] Using a pressurised coaxial needle with the exteriorfluid comprising a viscous polydimethylsiloxane solution and an aqueouscell suspension in the core, cells can be processed into scaffolds with nonoticeable loss in viability.

Electrospinning and melt/solution blowing can be combined in a processknown as electroblowing. This employs both electricity and a gas flow toaid fibre elongation and solidification. The experimental apparatus usesa similar spinneret to that in [Figure apparatus used for solution blow spinning or melt blowing], and in addition to the gas flowa potential difference is applied between the spinneret and the collector.This technique has been shown to have significant potential in medicalapplications: in 2014, Jiang et al. applied electroblowing in vitro and invivo to deliver a homogeneous and continuous layer of a medical glue tostop bleeding during liver resection. [31]

More recently, Balogh et al. prepared fibres of 2-hydroxypropyl-\ensuremath{\beta }-cyclodextrin loaded with sodium diclofenac. [32] They found that whenelectrospinning this system, very frequent clogging of the spinneretoccurred. Electroblowing overcame this issue and additionally allowedfaster flow rates to be used, increasing the amount of material thatcould be produced. However, the uniformity of the fibre products wascompromised, with more `beads-on-string' type morphology seen withthe blown products. In both cases, the fibres comprised amorphous soliddispersions with no crystalline drug evident. The electroblown fibresdissolved a little more slowly than those from electrospinning, but stillmuch more rapidly than a physical mixture of drug and cyclodextrin.

A subsequent study using Eudragit E and itraconazole (anantifungal active pharmaceutical ingredient) also found that a fasterflow rate could be used in blowing, but that the fibre products from thelatter had less regular morphologies. [33] Again, the drug was amorphouslydispersed in the fibres, and the dissolution profiles of the electrospun andelectroblown systems were very similar.

\subsubsection{Microfluidic forces}Microfluidic spinning is based on the use of micro (sub-millimetre)channels. A large number of these are located in a single microfluidicchip, and the rate and time of liquid expulsion from each channel are precisely controlled by computer. Microfluidic spinning can be coupled withelectrospinning, as described in detail in a recent review by Cheng et al. [51]While the productivity of microfluidic methods is a major challenge inscale-up, the technique offers the ability to generate fibres with a highlevel of complexity not easily achievable by electrospinning. For example,using a digitally programmed microfluidic flow, Kang et al. created functional microfibres with continuous spatiotemporal coding along thelength of the fibre. [52] The fibres contained varied chemical compositionsand topography, and localised bioactive agents.

Microfluidic spinning therefore has the potential to enable veryprecisely tuneable loading of different drugs into a single fibre, allowingfor programmable release in different parts of the body at differenttimes. The technique is beginning to be explored for drug deliveryapplications. However, the materials used in microfluidics are usuallyhydrogels (crosslinked polymer networks solvated with water). Theseoften have fast degradation rates and as a result can be unsuitable forthe extended release of drugs, particularly small molecules. To helpmitigate this problem, Chae et al. developed a microfluidic spinningmethod using an isopropyl alcohol sheath flow with an aqueousalginate core flow. [53] This innovation resulted in nanofibres made ofhighly ordered alginate molecules. Ahn et al. loaded ampicillin intoalginate fibres prepared in this manner, [54 and found that the orderedstructure delayed fibre degradation, allowing extended-release profileof ampicillin over 7 days.

\subsubsection{Mechanical force}

\subsection{Dispensing nozzle: Coaxial / Monoaxial: Single nozzle / Multinozzle}

\subsection{Nozzle-to-substrate distance}

\subsubsection{Near Field Spinning}Figure~\ref{f-fe28447572e9} describes a typical near-field electrospinning set-up \unskip~\cite{527120:12073538}. Two sub-techniques can be derived from electrospinning depending on the distance between the dispensing electrode and the collector. The process in which the electrospun jet can be controlled near the tip is called NFES or near-field electrospinning \unskip~\cite{527120:12033655}. Moreover, if the distance between the collector and the dispensing needle is greater, the configuration is known as FFES or far-field electrospinning \unskip~\cite{527120:12073581}.

\subsubsection{Far Field Spinning}In  a far field electrospinning the process continues ... Decreasing the jet diameter, the surface charge density increases and the resulting high repulsive forces split the jet into several smaller jets. The jet is seriously elongated by a bending and whipping processes caused by electrostatic repulsion initiated at small bends in the fiber, until it is finally deposited on the collector. Two types of collector may be used, either stationary or rotary collector.

@@
